---
title: "Parallel processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, error = F, fig.width = 9)
library(evsim)
library(dplyr)
library(lubridate)
library(ggplot2)
```

The functions `get_demand()` and `get_n_connections()` have time-series calculations that are repeated on a daily basis. A 24-hour period is considered an independent optimization window, thus the corresponding time-series calculations can be done in parallel. 

At the same time, in computer science, a common practice to reduce computation times is **parallel processing** when the CPU of the computer has multiple cores that can work independently. We can discover the number of cores in our computer CPU using function `parallel::detectCores()`:

```{r}
n_cores <- parallel::detectCores()
print(n_cores)
```

In this case, the CPU used is the Apple M4 Pro, which features 10 performance CPU cores (P-core) running at up to 4.5 GHz along with 4 efficient cores (E-core) running at up to 2.6 GHz. So, when `parallel::detectCores()` returns 14, macOS is exposing 14 logical cores — but not all of them are equal in speed.

## Time-series EV demand calculation

Package `evsim` provides a sample data set of California EV sessions from October 2018 to September 2021. If we filter sessions corresponding to 2019, we have >15.000 sessions, with an average of 70 charging sessions during working days.

```{r, echo = T, eval = F}
sessions_2019 <- evsim::california_ev_sessions_profiles %>% 
  filter(year(ConnectionStartDateTime) == 2019)
```

Let's use these real charging sessions to calculate their time-series demand with `get_demand()` function, using multiple cores but also different values of number of days (optimization windows):

```{r, echo = T, eval = F}
n_days_seq <- c(3, 7, 15, 30, 120, 365) # Days in a year
n_cores_seq <- c(1, seq(2, 10, 2)) # 10 performance cores (P-core)
cores_time <- tibble(
  days = rep(n_days_seq, each = length(n_cores_seq)),
  cores = rep(n_cores_seq, length(n_days_seq)), 
  time = 0
)

for (nd in n_days_seq) {
  message(nd, " days ---------------- ")
  sessions <- sessions_2019 %>%
    filter(date(ConnectionStartDateTime) <= dmy(01012019)+days(nd)) %>% 
    evsim::adapt_charging_features(time_resolution = 15)

  for (mcc in cores_time$cores[cores_time$days == nd]) {
    message(mcc, " cores - ", Sys.time())
    
    # Define daemons
    if (mcc > 1) {
      mirai::daemons(mcc)
    }
    # Execute calculations
    results <- system.time(
      sessions %>% 
        get_demand(by = "Profile", resolution = 15)
    )
    # Reset daemons
    mirai::daemons(0)
    
    cores_time$time[
      cores_time$cores == mcc & cores_time$days == nd
    ] <- as.numeric(results[3])
  }
}

# Adapt variables for a better plot
cores_time <- cores_time %>% 
    mutate(
      days = factor(
        paste(days, "days"), 
        levels = factor(paste(n_days_seq, "days"))
      ), 
      cores = factor(cores)
    )
```


```{r saved data, echo=F, eval = T}
# saveRDS(cores_time, "parallel_processing_data/cores_time.RDS")
cores_time <- readRDS("parallel_processing_data/cores_time.RDS")
```

The plot below shows that:

1. Parallel processing is worth it from more than 1-month data sets (at least for this case with ~70 charging sessions per day)
2. From 6 cores the reduction in computation time is not very relevant


```{r}
#| fig.alt: >
#|   Computation time for battery optimization using multiple cores
cores_time %>%
  ggplot(aes(x = cores, y = time)) +
  geom_col() + 
  facet_wrap(vars(days), scales = "free", nrow = 2) + 
  labs(x = "Number of cores", y = "Computation time (seconds)")
```

